---
title: "Coursera - Reproducible Research - Week 2- Project Assessment 1"
author: "Benoit de TARADE"
date: "6 janvier 2017"
output: html_document
---

# Install package if need and load it


```{r setup echo:FALSE}


if (!require("ggplot2")) {
  install.packages("ggplot2")
}


require("ggplot2")

```

# Upload, Unzip and read data

- Url used for download : https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip
- Output file activity.zip
- Readed file activity.csv

```{r fileUploadUnzip}
f <- download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip")
unzip("activity.zip", overwrite = TRUE)


## Read activity and features labels
activity <- read.csv("activity.csv")
```

# Histogram of the total number of steps taken each day


```{r totalPerDay}
# Calculate the total number of steps taken per day
totalSteps <- tapply(activity$steps, activity$date, FUN=sum, na.rm=TRUE)
# Histogram for the Total Steps
hist(totalSteps, main="Histogram of Total Steps", xlab="Total number of steps taken each day")
```



#Mean and median number of steps taken each day

```{r meanAndMedian}
# Calculate and report the mean and median of the total number of steps taken per day
meanVal <- mean(totalSteps, na.rm=TRUE)
medianVal <- median(totalSteps, na.rm=TRUE)

```

- The mean value for the # of steps taken is `r meanVal`
- The median value for the # of steps taken is `r medianVal`

# Time series plot of the average number of steps taken

```{r averageDailyActivity}
activity.avg <- aggregate(x=list(steps=activity$steps), by=list(interval=activity$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=activity.avg, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("Average # of steps taken")
```

# The 5-minute interval that, on average, contains the maximum number of steps

```{r maxSteps}
maxstep <- activity.avg[which.max(activity.avg$steps),]
```

- Maximum of Steps in 5-Minute Interval is `r maxstep`



# Code to describe and show a strategy for imputing missing data

```{r missingValue}
activity.missing <- sum(is.na(activity$steps))
```

- Total number of missing values `r activity.missing`


All of the missing values are filled in with mean value for that 5-minute interval.

Replace each missing value with the mean value of its 5-minute interval
```{r replaceMean}

activity.repl <- function(steps, interval) {
    rc <- NA
    if (!is.na(steps))
        rc <- c(steps)
    else
        rc <- (activity.avg[activity.avg$interval==interval, "steps"])
    return(rc)
}

activity.comp <- activity
activity.comp$steps <- mapply(activity.repl, activity.comp$steps, activity.comp$interval)
```


```{r drawHist2}
activity.steps <- tapply(activity.comp$steps, activity.comp$date, FUN=sum)
hist(activity.steps, main="Histogram of Total Steps - No NA's", xlab="Total number of steps taken each day")
activity.steps.mean <- mean(activity.steps)
activity.steps.median <- median(activity.steps)
```
The mean value for the # of total steps taken without NA's is `r activity.steps.mean`

The median value for the # of total steps taken without NA's is `r activity.steps.median`


# Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

## Build function to split weekdays and weekends

Beware!!! If your'nt in english settings there will be only one facet 'bad language'

```{r dayOfWeek}
weekDay  <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        return("bad language")
}

```

## Build plots

```{r stepsforWeek}
activity.comp$date <- as.Date(activity.comp$date)

activity.comp$weekend <- sapply(activity.comp$date, FUN=weekDay)


activity.avg <- aggregate(steps ~ interval + weekend, data=activity.comp, mean)
ggplot(activity.avg, aes(interval, steps)) + geom_line() + facet_grid(weekend ~ .) +
    xlab("5 Minute Interval") + ylab("# of steps")
```
